\chapter{Introduction}

In the last decades, the IoT industry has experienced an explosive growth, both in numbers of \textit{connected} devices and total industry worth. This trend has been closely followed by a rise in the types and amounts of cyberattacks that target these IoT devices, causing enormous economic, privacy and material losses to businesses, governments and people worldwide.\\

Cyberattacks on IoT devices have increased greatly in the past years \cite{forbes}, and this trend will only continue in the future. This, coupled with the fact that almost half of businesses claim not to be able to detect IoT security breaches in their devices \cite{gemalto}, reveals the need for the improvement of IoT device protection from cyberattacks and detection of malicious behaviour.\\

The aim of this project is two-fold, firstly provide a research of the current IoT system configurations, studying possible network configurations, communication protocols and attack surfaces that can affect the devices. Secondly, a method of detecting and mitigating the severity of attacks will be explained through the use of monitoring on IoT systems. Several theoretical real world use cases will be proposed to help the reader understand some potential attacks, consequences and monitoring solutions in IoT clusters.\\

To further prove this, a small scale demo will be implemented with the ESP32 development board and the FreeRTOS Operating System to demonstrate some aspect of the monitoring of the FreeRTOS running on the ESP32 development boards and how intelligent information can be extracted from the parameters that are monitored.\\

The remainder of the document is organized as follows; The background will consist of studying the IoT industry, Communication protocols, Network topologies, FreeRTOS  and ESP-IDF and Attacks on IoT systems. After this several Use cases will be proposed and the demo implementation will be explained. Lastly some results and conclusions will be analyzed, which together with the bibliography will mark the end of this report. A user guide on how to replicate the demo is included as an appendix. 





